# Data 

## Sources
The data is published on the NBA official website, in specific, you can:

|==>Find 2021-2022 regular season data through this link: *[regular season](https://www.nba.com/stats/teams/traditional?Season=2021-22)

|==>Find 2021-2022 playoff data through this link: *[playoff](https://www.nba.com/stats/teams/traditional?Season=2021-22&SeasonType=Playoffs)

|==>Find final champion data trough this link: *[final champion](https://www.nba.com/stats/team/1610612744/traditional)

## Cleaning / transformation

```{r, include=FALSE}
# this prevents package loading message from appearing in the rendered version of your problem set
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```


First thing first, we load the data
```{r}
library(rvest)
leader1 = read_html('./docs/data/Teams Traditional _ Stats _ NBA.com.html') %>%
    html_nodes("table") %>%
    html_table(header = T)

leader2 = read_html('./docs/data/Teams Traditional _ Stats _ NBA2.com.html') %>%
    html_nodes("table") %>%
    html_table(header = T)

ori_regular <- leader1[[3]][,1:28]
ori_playoff <- leader2[[3]][,1:28]
(head(ori_regular))
(head(ori_playoff))

```

Second, we do some cleaning.

```{r}
#Rename the first column as the original data do not have a column name.
names(ori_regular)[1] <- "Rank"
names(ori_playoff)[1] <- "Rank"
```

Third, we do some transformation.
```{r}
ori_regular$`WIN%`<-ori_regular$`WIN%`*100
ori_playoff$`WIN%`<-ori_playoff$`WIN%`*100
```


We have to admit that because those data is collected by NBA official with most advanced equipment, so it is quite clean and well structured. By carefully checking every rows and columns, we finally find some uncleaned data: a missing column name and not clear percentage numbers. 

For further transformation, we can consider using standard scaler or log scaler for those numerical features, but this final project so far only focus on data visualization and analyzation, not about training a machine learning model, so we do not have to further scale those data. For data cleaning and transformation step, it is enough so far.


## Missing value analysis

Now we go into deep to check the missing value.
```{r}
# we check if there is missing values in each column
print('Missing values count in regular season:')
(colSums(is.na(ori_regular)))
print('Missing values count in playoff:')
(colSums(is.na(ori_playoff)))
```


Good thing is there is no missing values inside, the dataset is somehow 'clean' and 'beautiful' enough.
