# Data 

## Sources
The data used in this analysis is collected and published on the NBA official website, the following links direct to the raw data.

|==> Data of 2021-2022 regular season: *[regular season](https://www.nba.com/stats/teams/traditional?Season=2021-22)

|==> Data of 2021-2022 playoff: *[playoff](https://www.nba.com/stats/teams/traditional?Season=2021-22&SeasonType=Playoffs)

|==> Data of Final Champions: *[final champion](https://www.nba.com/stats/team/1610612744/traditional)

## Cleaning / transformation

```{r, include=FALSE}
# this prevents package loading message from appearing in the rendered version of your problem set
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```


First of all, we will load the data for preprocessing.

```{r}
library(rvest)
leader1 = read_html('./docs/data/Teams Traditional _ Stats _ NBA.com.html') %>%
    html_nodes("table") %>%
    html_table(header = T)

leader2 = read_html('./docs/data/Teams Traditional _ Stats _ NBA2.com.html') %>%
    html_nodes("table") %>%
    html_table(header = T)

ori_regular <- leader1[[3]][,1:28]
ori_playoff <- leader2[[3]][,1:28]
(head(ori_regular))
(head(ori_playoff))

```

Secondly, we will perform data cleaning.

```{r}
#Rename the first column as the original data do not have a column name.
names(ori_regular)[1] <- "Rank"
names(ori_playoff)[1] <- "Rank"
```

Thirdly, we do some transformation.
```{r}
ori_regular$`WIN%`<-ori_regular$`WIN%`*100
ori_playoff$`WIN%`<-ori_playoff$`WIN%`*100
```


We noticed that the data is mostly clean and well structured, since it was collected by NBA officially. By scrutinizing the whole data, the major data cleaning we performed were fixing column names and unclear percentage values. 

For further transformations, we would consider using standard scaler or log scaler for numerical features. Since the main goal for final project is about data visualization and analyzation, further scaling or tranforming of the data is not necessary. 


## Missing value analysis

In this step, we will perform the missing value analysis.
```{r}
# we check if there is missing values in each column
print('Missing values count in regular season:')
(colSums(is.na(ori_regular)))
print('Missing values count in playoff:')
(colSums(is.na(ori_playoff)))
```

Using library **redav** to create two missing values visualization.

We can very clearly see from those two graphs that there is no missing value in both regular season and playoffs.

```{r}
library(redav)

plot_missing(ori_regular, percent = FALSE)
plot_missing(ori_playoff, percent = FALSE)

```

What's more, I create a heapmap to show the missing values. Still, we can see from those two heatmaps clearly that there is no missing values.

```{r}
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2)
library(forcats)
# For playoff 
tidy_playoff <- ori_playoff %>% 
    #rownames_to_column("id") %>% 
    gather(key, value, -Team) %>% 
    mutate(missing = ifelse(is.na(value), "yes", "no"))

ggplot(tidy_playoff, aes(x = key, y = fct_rev(Team), fill = missing)) +
  geom_tile(color = "white") + 
  ggtitle("playoff with NAs added") +
  ylab('') + 
  scale_fill_viridis_d() + # discrete scale
  theme_bw()

# For regular season
tidy_regular <- ori_regular %>% 
    #rownames_to_column("id") %>% 
    gather(key, value, -Team) %>% 
    mutate(missing = ifelse(is.na(value), "yes", "no"))

ggplot(tidy_regular, aes(x = key, y = fct_rev(Team), fill = missing)) +
  geom_tile(color = "white") + 
  ggtitle("playoff with NAs added") +
  ylab('') + 
  scale_fill_viridis_d() + # discrete scale
  theme_bw()
```



From the above tables we can see that there exist none missing values in our data. Hence, no filling of missing data is needed.
