--- 
title: "NBA 2021 Season Analysis"
author: "Mouwei Lin (lm3756); Frank Li (zl3204); Linhao Yu (ly2590)"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction
The National Basketball Association (NBA) is a professional basketball league in North America. The league is composed of 30 teams (29 in the United States and 1 in Canada) and is one of the major professional sports leagues in the United States and Canada. It is the premier men's professional basketball league in the world.

As NBA attract millions of viewers each year, it is definitely one of the most famous league in the world now. That also means there are plenty of data collected and ready to be analyzed. In this final project, we will focus on 2021 to 2022 NBA season, try to use data to persuade people why some teams win the game and why they are extraordinary this year.




<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Proposal

## Research topic
NBA is one of the most famous sports leagues in the world. In this final project, we will focus on data analysis of NBA regular season and playoffs in season 2021-2022. 

In regular season, we will analyze each team's performance base on conference. Performance will include: field goals made(FGM), 3 point field goals made(3PM), free throws made(FTM), assists(AST), offensive rebounds(OREB), defensive rebounds(DREB), blocks(BLK) and steals(STL). Among these features, FGM, 3PM, FTM, AST and OREB are categorized to be offensive statistics since they are related to a team's offensive performance. DREB, BLK and STL are considered as defensive statistics since they reveal the defense level of a team. We will use some visualization techniques to check different conference's styles based on these statistics and determine if those styles significantly influence each team's victory rate.

In playoffs, firstly, we will do similar analyze like regular season. Then we will discuss why those teams perform better than those that can not enter the playoffs. Finally, we will analyze the teams played in NBA finals in 2021, to see their performance in every round.

## Data availability
In this *[link](https://www.nba.com/stats/help/faq), NBA official website has clearly explained how they collected and distributed those data. In conclusion, the information will be like following:

|==>**How data collected**: according to NBA official's explanation, all those data, which include:

-GP: Games Played

-W: Wins

-L: Losses

-WIN%: Win Percentage

-MIN: Minutes Played

-PTS: Points

-FGM: Field Goals Made

-FGA: Field Goals Attempted

-FG.: Field Goal Percentage

-3PM: 3 Point Field Goals Made

-3PA: 3 Point Field Goals Attempted

-3P.: 3 Point Field Goal Percentage

-FTM: Free Throws Made

-FTA: Free Throws Attempted

-FT%: Free Throw Percentage

-OREB: Offensive Rebounds

-DREB: Defensive Rebounds

-REB: Rebounds

-AST: Assists

-TOV: Turnovers

-STL: Steals

-BLK: Blocks

-BLKA: Blocks Against

-PF: Personal Fouls

-PFD: Personal Fouls Drawn

-+/-: Plus-Minus

-Enter: Weather or not This Team Enter Playoff

were collected by Using cameras installed in the catwalks of every NBA arena, Second Spectrum software tracks the movements of every player on the court and the basketball 25 times per second. You can go this *[page](https://www.nba.com/stats/players/catch-shoot) for more information. 
The tools to collect those data are most advanced and accurate. 
All the advanced stats from NBA website can go back to the 1996-97 season; however their base stats go back to the inaugural 1946-47 season. Every box score has been digitized and those go back to the 1946-47 season as well. Their lineup data goes back to 2008. Certain stats were not recorded from the beginning of the NBA. Here’s a list of years when some base stats began being recorded:

-Rebounds: 1950-1951	

-Minutes: 1951-1952	

-Games Started: 1970-1971

-Steals: 1973-1974	

-Blocks: 1973-1974	

-Off Rebounds: 1973-1974

-Def Rebounds: 1973-1974	

-Turnovers: 1977-1978	

-3 Point Field Goals: 1979-1980

|==>**Who collect the data**: usually all the data was collected by NBA official. 

|==>**Format of data**: because there is no cvs file provided, we use *revest* , a R package to read the data directly from html. So it will be several data frames and all features are numeric data types.

|==>**Frequency of updates**: according to NBA website's explanation, a player or team’s base statistics are updated simultaneously with the game and in real time. The advanced stats are updated between 10-15 minutes after a game finishes. Moreover, because of the data's own properties, it usually has very few updates once it is published.

|==>**Sources:**
We will post the sources' links here again:

|==>Find 2021-2022 regular season data through this link: *[regular season](https://www.nba.com/stats/teams/traditional?Season=2021-22)

|==>Find 2021-2022 playoff data through this link: *[playoff](https://www.nba.com/stats/teams/traditional?Season=2021-22&SeasonType=Playoffs)

|==>Find final champion data trough this link: *[final champion](https://www.nba.com/stats/team/1610612744/traditional)




<!--chapter:end:proposal.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data 

## Sources
The data used in this analysis is collected and published on the NBA official website, the following links direct to the raw data.

|==> Data of 2021-2022 regular season: *[regular season](https://www.nba.com/stats/teams/traditional?Season=2021-22)

|==> Data of 2021-2022 playoff: *[playoff](https://www.nba.com/stats/teams/traditional?Season=2021-22&SeasonType=Playoffs)

|==> Data of Final Champions: *[final champion](https://www.nba.com/stats/team/1610612744/traditional)

## Cleaning / transformation

```{r, include=FALSE}
# this prevents package loading message from appearing in the rendered version of your problem set
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```


First of all, we will load the data for preprocessing.

```{r}
library(rvest)
leader1 = read_html('./docs/data/Teams Traditional _ Stats _ NBA.com.html') %>%
    html_nodes("table") %>%
    html_table(header = T)

leader2 = read_html('./docs/data/Teams Traditional _ Stats _ NBA2.com.html') %>%
    html_nodes("table") %>%
    html_table(header = T)

ori_regular <- leader1[[3]][,1:28]
ori_playoff <- leader2[[3]][,1:28]
(head(ori_regular))
(head(ori_playoff))

```

Secondly, we will perform data cleaning.

```{r}
#Rename the first column as the original data do not have a column name.
names(ori_regular)[1] <- "Rank"
names(ori_playoff)[1] <- "Rank"

#Factor the Team and Rand column, make it easy to analyze.
ori_playoff$Team = as.factor(ori_playoff$Team)
ori_playoff$Rank = as.factor(ori_playoff$Rank)
ori_regular$Team = as.factor(ori_regular$Team)
ori_regular$Rank = as.factor(ori_regular$Rank)
```


Thirdly, we do some transformation, we add a new column **Enter** in regular season, if Enter is True, it means this team enter the playoff, otherwise it does not enter playoff.
```{r}
ori_regular$`WIN%`<-ori_regular$`WIN%`*100
ori_playoff$`WIN%`<-ori_playoff$`WIN%`*100
ori_regular$Enter = (ori_regular$Team %in% ori_playoff$Team)
```


We noticed that the data is mostly clean and well structured, since it was collected by NBA officially. By scrutinizing the whole data, the major data cleaning we performed were fixing column names and unclear percentage values. 

For further transformations, we would consider using standard scaler or log scaler for numerical features. Since the main goal for final project is about data visualization and analyzation, further scaling or tranforming of the data is not necessary. 


## Missing value analysis

In this step, we will perform the missing value analysis.
```{r}
# we check if there is missing values in each column
print('Missing values count in regular season:')
(colSums(is.na(ori_regular)))
print('Missing values count in playoff:')
(colSums(is.na(ori_playoff)))
```

Using library **redav** to create two missing values visualization.

We can very clearly see from those two graphs that there is no missing value in both regular season and playoffs.

```{r}
library(redav)

plot_missing(ori_regular, percent = FALSE)
plot_missing(ori_playoff, percent = FALSE)

```

What's more, I create a heapmap to show the missing values. Still, we can see from those two heatmaps clearly that there is no missing values.

```{r}
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2)
library(forcats)
# For playoff 
tidy_playoff <- ori_playoff %>% 
    #rownames_to_column("id") %>% 
    gather(key, value, -Team) %>% 
    mutate(missing = ifelse(is.na(value), "yes", "no"))

ggplot(tidy_playoff, aes(x = key, y = fct_rev(Team), fill = missing)) +
  geom_tile(color = "white") + 
  ggtitle("playoff with NAs added") +
  ylab('') + 
  scale_fill_viridis_d() + # discrete scale
  theme_bw()

# For regular season
tidy_regular <- ori_regular %>% 
    #rownames_to_column("id") %>% 
    gather(key, value, -Team) %>% 
    mutate(missing = ifelse(is.na(value), "yes", "no"))

ggplot(tidy_regular, aes(x = key, y = fct_rev(Team), fill = missing)) +
  geom_tile(color = "white") + 
  ggtitle("playoff with NAs added") +
  ylab('') + 
  scale_fill_viridis_d() + # discrete scale
  theme_bw()
```



From the above tables we can see that there exist none missing values in our data. Hence, no filling of missing data is needed.


Last thing we do is storing above data we cleaned, for the next step.
```{r}
write.csv(ori_playoff,"./docs/data/playoff.csv", row.names = FALSE)
write.csv(ori_regular,"./docs/data/regular.csv", row.names = FALSE)
```


<!--chapter:end:data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

```{r, include=FALSE}
# this prevents package loading message from appearing in the rendered version of your problem set
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

First thing first, that's load the data we cleaned on **data** page.

```{r}
#Read cleaned data
playoff<-read.csv(file = './docs/data/playoff.csv')
regular<-read.csv(file = './docs/data/regular.csv')

#Factor Team and Rank
playoff$Team = as.factor(playoff$Team)
playoff$Rank = as.factor(playoff$Rank)
regular$Team = as.factor(regular$Team)
regular$Rank = as.factor(regular$Rank)
```


##  Points with Rank in Playoff and Regular Season

```{r}
library(ggplot2)

#plot scatter plot with smooth line in playoff about points
g_playoff <- ggplot(data = playoff,aes(x=Rank, y=PTS))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank), y=PTS))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Average Points in Playoff')+
  ylab('Points')
g_playoff

#plot scatter plot with smooth line in regular about points
g_regular <- ggplot(data = regular,aes(x=Rank, y=PTS))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=PTS))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Average Points in Regular')+
  ylab('Points')
g_regular
```

We first analyze the teams that enter the playoff because they have higher commercial values in real world. We can find that once teams enter playoff, their average points are very close to each other, on the other word, their skills are quite similar, that's why playoff in NBA is usually more interesting than the regular season, because every team has the chance to win. In conclusion, if we only look at the points each team get on average during playoff, they are very similar with each other. Only the bottom few perform a little worse.


While we look back in regular season, things are quite different. There is a very clear difference in the points between the teams who did or did not enter playoff. As we can see that first 16 teams (who entered the playoff) still have very similar points gain, while in last 14 teams (who did not enter playoff), their points gain dropped with the ranking. On the other word, we can say that the NBA's official competition system is very reasonable. It screens out strong teams and weak teams very well, and the playoffs are even more exciting because of that.



##  Point Field Goals with Rank in Playoff and Regular Season

```{r}
library(patchwork)

#plot scatter plot with smooth line in playoff about 3 points
g_playoff <- ggplot(data = playoff,aes(x=Rank, y=X3PM))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank), y=X3PM))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('3 Point Field Goals Made and 3 Point\nField Goals Percentage in Playoff')+
  ylab('3 Point Field Goals Made')
g_playoff2 <- ggplot(data = playoff,aes(x=Rank, y=X3P.))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank), y=X3P.))+
  theme(axis.text.x = element_text(angle=90))+
  ylab('3 Point Field Goals Percentage')

g_playoff / g_playoff2 + plot_layout(guides = 'collect')

#plot scatter plot with smooth line in regular about 3 points
g_regular <- ggplot(data = regular,aes(x=Rank, y=X3PM))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=X3PM))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('3 Point Field Goals Made and 3 Point\nField Goals Percentage in Regular')+
  ylab('3 Point Field Goals Made')
g_regular2 <- ggplot(data = regular,aes(x=Rank, y=X3P.))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=X3P.))+
  theme(axis.text.x = element_text(angle=90))+
  ylab('3 Point Field Goals Percentage')

g_regular / g_regular2 + plot_layout(guides = 'collect')
```

We now analyze the 3 Point Field performance. We first start with playoff. The trend in playoff about 3 Point is quite similar to total points, their average 3 point goals have very little differences. However, interestingly, when we check the 3 point goals percentage, we can see that the last team has the highest percentage, which means it has highest three-point field goal percentage. This can be explained by the level gap, or on the other word, although it looks like every team in playoff has very similar points goals, because of the level gap, weak teams have less offensive ability on the basketball court, the way to catch up with the strong teams is to shoot more three points. But unfortunately, this is a good strategy when fight with weak teams but not usually work with strong teams. That's why it ranked 16 in playoff.


When we look at regular season, will can find that strong teams (who enter playoff) have both the high 3 point goals made and 3 point goals percentage. Top ranked teams among weak teams (do not enter playoff) have both low 3 point goals made and 3 point goals percentage. However, those last few teams have a very high 3 point goals made but lowest 3 point goals percentage. It can also be explained by the level gap. Strong teams are always strong, they can make 3 point easier and offensive efficiently. Top ranked teams among weak teams may have good offensive abilities, but they lack the 3 points goals to open up the gap in order to win, that's why they lose. While the last few teams, they are desperately to shoot more 3 points in order to get the points, they are weak in offensive so that they have to shoot 3 points, but they also not good at it, that's why their percentages are so low either.




##  Field Goal with Rank in Playoff and Regular Season

```{r}
#plot scatter plot with smooth line in playoff about field goals
g_playoff <- ggplot(data = playoff,aes(x=Rank, y=FGM))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank)))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Field Goals Made and Field Goals \nPercentage in Playoff')+
  ylab('Field Goals Made')
g_playoff2 <- ggplot(data = playoff,aes(x=Rank, y=FG.))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank), y=FG.))+
  theme(axis.text.x = element_text(angle=90))+
  ylab('Field Goals Percentage')

g_playoff / g_playoff2 + plot_layout(guides = 'collect')

#plot scatter plot with smooth line in regular about field goals
g_regular <- ggplot(data = regular,aes(x=Rank, y=FGM))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=FGM))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Field Goals Made and Field Goals \nPercentage in Regular')+
  ylab('Field Goals Made')
g_regular2 <- ggplot(data = regular,aes(x=Rank, y=FG.))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=FG.))+
  theme(axis.text.x = element_text(angle=90))+
  ylab('Field Goals Percentage')

g_regular / g_regular2 + plot_layout(guides = 'collect')
```

Then we look at the general field goals performance. This is the most critical factor that reflects a team's performance because most of the time, a team win points through general field goals.


In playoff, things are a bit different now, although the trend is very weak, field goals made drops slightly as rank drops while in field goals percentage, they are quite similar. That means each team's shooting percentage is similar, all of them can shoot very good. However, stronger teams have more opportunities to shoot because of their stronger offensive abilities. That's why they can win. And we can conclude that although each team in playoff is strong, they still have subtle level gap.


In regular season this trend becomes very notable. Strong teams always have higher field goals made and field goals percentage while weak teams have lower. They have a very clear level gap.

##  Free Throw with Rank in Playoff and Regular Season

```{r}
#plot scatter plot with smooth line in playoff about Free Throw 
g_playoff <- ggplot(data = playoff,aes(x=Rank, y=FTM))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank)))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Free Throws Made and Free Throws \nPercentage in Playoff')+
  ylab('Free Throws Made')
g_playoff2 <- ggplot(data = playoff,aes(x=Rank, y=FT.))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank), y=FT.))+
  theme(axis.text.x = element_text(angle=90))+
  ylab('Free Throws Percentage')

g_playoff / g_playoff2 + plot_layout(guides = 'collect')

#plot scatter plot with smooth line in regular about Free Throw 
g_regular <- ggplot(data = regular,aes(x=Rank, y=FTM))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=FTM))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Free Throws Made and Free Throws \nPercentage in Regular')+
  ylab('Free Throws Made')
g_regular2 <- ggplot(data = regular,aes(x=Rank, y=FT.))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=FT.))+
  theme(axis.text.x = element_text(angle=90))+
  ylab('Free Throws Percentage')

g_regular / g_regular2 + plot_layout(guides = 'collect')
```

Now we will analyze the effect of Free Throws. Free throws are generally awarded after a foul on the shooter by the opposing team, analogous to penalty shots in other team sports. Free throws are also awarded in other situations, including technical fouls, and when the fouling team has entered the bonus/penalty situation. Also, depending on the situation, a player may be awarded between one and three free throws. Each successful free throw is worth one point.

Intuitively thinking, Free throws serve as a supporting factor for winning, not the key factor, since Free Throws only worth one point each, and the accuracy depends on the player heavily. We should not expect more Free Throws resulting in higher rank.

As expected, in Playoff the Free Throws made does not lead to a higher ranking. In fact, the team in the first place even has a low amount of Free Throws made. 

This happened again in the analysis in Regular Season, in Regular Season the Free Throws made also does not lead to a higher ranking. In fact, the team in the first place even has a low amount of Free Throws made in general.

##  Rebounds made and Rank in Playoff and Regular Season

```{r}
#plot scatter plot with smooth line in playoff about Rebounds Made
ggplot(data = playoff,aes(x=Rank, y=REB))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank)))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Rebounds Made in Playoff')+
  ylab('Rebounds Made')



#plot scatter plot with smooth line in regular about Rebounds Made
ggplot(data = regular,aes(x=Rank, y=REB))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=REB))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Rebounds Made in Regular')+
  ylab('Rebounds Made')


```

Then we will analyze the effect of Rebounds made. A rebound, or board, is a statistic awarded to a player who retrieves the ball after a missed field goal or free throw.Rebounds in basketball are a routine part in the game; if a shot is successfully made possession of the ball will change, otherwise the rebound allows the defensive team to take possession. Rebounds are also given to a player who tips in a missed shot on his team's offensive end. A rebound can be grabbed by either an offensive player or a defensive player.

Intuitively thinking, Rebounds made serve as a key factor for winning, since each rebound means the opportunity for a team to be ofensive, and have the change to score points. We should expect more rebounds being made resulting in higher rank.

As expected, in Playoff the Rebounds made does lead to a higher ranking. In fact, the team in the first place even has a high amount of rebounds made. 

This happened again in the analysis in Regular Season more obviously, in Regular Season the rebounds made also leads to a higher ranking. In fact, the team in the first place even has a how amount of rebounds made in general.


##  Personal Fouls made and Rank in Playoff and Regular Season

```{r}
#plot scatter plot with smooth line in playoff about Personal Fouls
ggplot(data = playoff,aes(x=Rank, y=PF))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team))+
  geom_smooth(aes(x=as.integer(Rank)))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Personal Fouls Made in Playoff')+
  ylab('Personal Fouls Made')



#plot scatter plot with smooth line in regular about Personal Fouls
ggplot(data = regular,aes(x=Rank, y=PF))+
  geom_point(size=3.0,stroke=0.1,aes(color=Team,shape=Enter))+
  geom_smooth(aes(x=as.integer(Rank), y=PF))+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle('Personal Fouls Made in Regular')+
  ylab('Personal Fouls Made')


```

Finally we will analyze the effect of Personal Fouls. A personal foul is a breach of the rules that concerns illegal personal contact with an opponent. It is the most common type of foul in basketball. A player fouls out on reaching a limit on personal fouls for the game and is disqualified from participation in the remainder of the game. Players routinely initiate illegal contact to purposely affect the play, hoping it is seen as too minor to be ruled a foul. The threshold is subjective and varies among officials and from game to game. Most contact fouls are not regarded as unsportsmanlike.

Intuitively thinking, Personal Fouls should have a negative correlation with ranking, since Personal Fouls will provide more attacking chance for the opposite team, and the player might have to retreat. We should expect less Personal Fouls in higher rank.

As expected, in Playoff less Personal Fouls does lead to a higher ranking. In fact, the team in the first place even has a low amount of Personal Fouls. 

This does not happened again in the analysis in Regular Season, in Regular Season less Personal Fouls does not lead to a higher ranking from the plot. One of the reasons could be the skill differences are to large between teams. Personal Fouls factor is not strong enough to secure a high rank, although more attacking opportunities were provided.


<!--chapter:end:results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component
In this part, we use two interactive lollipop charts to visualize different statistics of NBA teams in regular season and playoff. In the charts,teams are ordered based on their winning percentage, and we use light blue dots to represent teams from western conference while dark blue represents eastern conference. 

There are 8 statistics to choose from: field goals made(FGM), 3 point field goals made(3PM), free throws made(FTM), assists(AST), offensive rebounds(OREB), defensive rebounds(DREB), blocks(BLK) and steals(STL). One can use the buttons at the top to select an interested statistics and see the change. 


For the interactive component part, please click on this link below to access:

**[Regular season](interaction_regular.html)**

**[Playoff](interaction_playoff.html)**

<!--chapter:end:interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion

We have obtained a broad view on the NBA's data, and perfermed several analysis. In this final project, our focus is on data of NBA regular season and playoffs in season 2021-2022. The major goal for our project is determining the relationship between the factors, and tried to explore which of them are keys factors for a team to win.

In regular season, we have analyzed each team’s performance base on conference. Performance includes: field goals made(FGM), 3 point field goals made(3PM), free throws made(FTM), assists(AST), offensive rebounds(OREB), defensive rebounds(DREB), blocks(BLK) and steals(STL). Among these features, FGM, 3PM, FTM, AST and OREB are categorized to be offensive statistics since they are related to a team’s offensive performance. DREB, BLK and STL are considered as defensive statistics since they reveal the defense level of a team. In playoffs, we did similar analyze like regular season. Then we discussed why those teams perform better than those that can not enter the playoffs.

As from the result sections, we obtained several conclustions from the analysis:

**1.** We found that strong teams (who enter playoff) have both the high 3 point goals made and 3 point goals percentage. Top ranked teams among weak teams (do not enter playoff) have both low 3 point goals made and 3 point goals percentage. However, those last few teams have a very high 3 point goals made but lowest 3 point goals percentage. It can be explained by the level gap. Strong teams are always strong, they can make 3 point easier and offensive efficiently. Top ranked teams among weak teams may have good offensive abilities, but they lack the 3 points goals to open up the gap in order to win, that’s why they lose.

**2.** For the general field goals performance. This is the most critical factor that reflects a team’s performance because most of the time, a team win points through general field goals. In playoff, although the trend is very weak, field goals made drops slightly as rank drops while in field goals percentage, they are quite similar. That means each team’s shooting percentage is similar, all of them can shoot very good. However, stronger teams have more opportunities to shoot because of their stronger offensive abilities. That’s why they can win. And we can conclude that although each team in playoff is strong, they still have subtle level gap. In regular season this trend becomes very notable. Strong teams always have higher field goals made and field goals percentage while weak teams have lower. They have a very clear level gap.

**3.** For the effect of Free throws, Intuitively thinking, Free throws serve as a supporting factor for winning, not the key factor, since Free Throws only worth one point each, and the accuracy depends on the player heavily. We should not expect more Free Throws resulting in higher rank. As expected, in Playoff the Free Throws made does not lead to a higher ranking. In fact, the team in the first place even has a low amount of Free Throws made. This happened again in the analysis in Regular Season, in Regular Season the Free Throws made also does not lead to a higher ranking. In fact, the team in the first place even has a low amount of Free Throws made in general.

**4.** For the effect of Rebounds, Intuitively thinking, Rebounds made serve as a key factor for winning, since each rebound means the opportunity for a team to be ofensive, and have the change to score points. We should expect more rebounds being made resulting in higher rank. As expected, in Playoff the Rebounds made does lead to a higher ranking. In fact, the team in the first place even has a high amount of rebounds made. This happened again in the analysis in Regular Season more obviously, in Regular Season the rebounds made also leads to a higher ranking. In fact, the team in the first place even has a how amount of rebounds made in general.

**5.** For the effect of Personal Fouls, Intuitively thinking, Personal Fouls should have a negative correlation with ranking, since Personal Fouls will provide more attacking chance for the opposite team, and the player might have to retreat. We should expect less Personal Fouls in higher rank. As expected, in Playoff less Personal Fouls does lead to a higher ranking. In fact, the team in the first place even has a low amount of Personal Fouls. This does not happened again in the analysis in Regular Season, in Regular Season less Personal Fouls does not lead to a higher ranking from the plot. One of the reasons could be the skill differences are to large between teams. Personal Fouls factor is not strong enough to secure a high rank, although more attacking opportunities were provided.

All the above conclusion seems to be reasonable for determining the path for the competition of each team, in the basketball perspective:

<!--chapter:end:conclusion.Rmd-->

